---
title: "Homework 6"
author: "Connie Zhang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
```

## Question 1
```{r - Question 1 - data cleaning}
birth_data = read_csv("./data/birthweight.csv") %>% 
janitor::clean_names() %>%
mutate (
  babysex = recode(babysex, "1" = "male", "2" = "female"),
  frace = recode(frace,
                        "1" = "white",
                        "2" = "black",
                        "3" = "asian",
                        "4" = "puetro rican",
                        "8" = "other",
                        "9" = "unknown"),
         mrace = recode(mrace,
                        "1" = "white",
                        "2" = "black",
                        "3" = "asian",
                        "4" = "puetro rican",
                        "8" = "other"),
           malform = recode(malform, 
                          "0" = "absent",
                          "1" = "present"
                          )) 
birth_data


#checking for missing values 
sum(!complete.cases(birth_data))

```

* There are no missing values in the data, as shown by the `0` value output. 

```{r - regression modeling}
birthmodel_fit = lm(bwt ~ 
                      gaweeks + 
                      delwt + 
                      fincome + 
                      smoken, 
                    data = birth_data) 

summary(birthmodel_fit) 
```

* These variables were chosen as well-known and regarded predictors of birthweight for newborns, incluidng both biological,, health, and socioeconomic factors. 

```{r - fitted vs residual}
birth_data %>%
  
  modelr::add_residuals(birthmodel_fit) %>%
  modelr::add_predictions(birthmodel_fit) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "green") + 
  labs(
        title = "Residuals vs. Predicted Values on a Hypothesized Model")

```

* The majority of residuals fall around the value of 0 and show constant variance. A few show as extremely high or low residuals, which could be outliers or data entry errors.  